Ad-hoc referential expressions for novel shapes in 4-5 year old children

### 2023-10-10 write-ups/CAMP6/abstract.tex
post mike comments

% Insert title, authors, affiliations, and e-mail address in the next three lines:

\title{Ad-hoc referential expressions for novel shapes in 4-5 year old children}
\authors{Veronica Boyce, Ilaria Chen, Bobby Sparks, Malia Perez, Michael C. Frank} 
\email{vboyce@stanford.edu;  Stanford University}
\newline
%

@@

As children acquire language, they must master many communication skills such as using metaphor and tracking the knowledge state of their partner. Adults demonstrate these communicative abilities in iterated reference games, where they describe novel figures to partners with high accuracy. Over repetitions, description length reduces as pairs come to share nicknames for the figures (Clark \& Wilkes-Gibbs 1986, Hawkins et al. 2020). This paradigm is commonly used with adults, but little used with pre-literate children.

In  Glucksberg et al. (1966), 5 year old children were not able to successfully communicate with each other about novel figures in a more complicated paradigm involving stacking blocks. However, children as young as 4 were successful at a two-choice, child-friendly iterated reference game when playing with their parent (Leung et al., 2020). Here, we test child-child dyads using a simplified version of Leung et al. 2020 to determine if children can communicate with each other and create partner-specific nicknames for novel figures in a simple iterated paradigm.

\textbf{Methods:} Pairs of children played a matching game on tablets. On each trial, the ``speaker'' child would see two images, one in a box, and describe the boxed image (orally) to their partner (Fig 1). The ``listener'' child saw the same two images and was tasked with helping Smurfy (a stuffy) guess the matching target image by touching it on the screen. Children alternated roles trial by trial; the children passed Smurfy back and forth to track their roles. The 4 target images were a subset of stimuli from Leung et al. 2020; distractors were drawn from the set of other targets. Children completed 2 practice trials with familiar shapes, and then 12 target trials (3 blocks of the 4 target images; Fig 1). We recruited 4-5 year old children from a university preschool; 20 pairs completed at least 8 out of 12 target trials and were included in the analysis. Utterances were transcribed from video recording using Whisper and then hand-checked. 
% Results

\textbf{Results:} 
Children uses a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page). Children were 84\% accurate at selecting the target image, and accuracy barely increased over time ($\beta=$0.13 [-.04, .31]; Fig 2A). Children's speed to select a target increased over time ($\beta=$-1.24 [-1.89, -.59]; Fig 2B). Children generally produced short descriptions, and the length of descriptions tended to slightly increase over time ($\beta=$0.11 [-.03, .25]; Fig 2C).

@@

>>>>

% Insert title, authors, affiliations, and e-mail address in the next three lines:
\noindent\title{4-5 year old children can successfully communicate using ad-hoc referential expressions}
\authors{Veronica Boyce, Ilaria Chen, Bobby Sparks, Malia Perez, Michael C. Frank} 
\email{vboyce@stanford.edu;  Stanford University}
\newline
%

@@

As children acquire language, they must master many communication skills such as using metaphor and tracking the knowledge state of their partner. Adults demonstrate these communicative abilities in iterated reference games, where they describe novel figures to partners with high accuracy. Over repetitions, description length reduces as pairs come to share nicknames for the figures (Clark \& Wilkes-Gibbs 1986, Hawkins et al. 2020). This paradigm is commonly used with adults, but the consensus has been that tracking ad-hoc references in this way is too difficult for young children (in part due to notable failures in the literature; Glucksberg et al., 1966). We revisit this question, using a simplified method adapted from Leung et al. 2020.

\bigskip
\textbf{Methods:} Pairs of children played a matching game on tablets. On each trial, the ``speaker'' child would see two images, one in a box, and describe the boxed image (orally) to their partner (Fig 1). The ``listener'' child saw the same two images and was tasked with helping Smurfy (a stuffy) guess the matching target image by touching it on the screen. Children alternated roles trial by trial; the children passed Smurfy back and forth to track their roles. The 4 target images were a subset of stimuli from Leung et al. 2020; distractors were drawn from the set of other targets. Children completed 2 practice trials with familiar shapes, and then 12 target trials (3 blocks of the 4 target images; Fig 1). We recruited 4-5 year old children from a university preschool; 20 pairs completed at least 8 out of 12 target trials and were included in the analysis. Utterances were transcribed from video recording using Whisper and then hand-checked. 
% Results

\bigskip

\textbf{Results:} 
Children uses a variety of different referential expressions to successfully convey the target to their partner (see Examples on next page). Children were 84\% accurate at selecting the target image, and accuracy was relatively stable over time ($\beta=$0.13 [-.04, .31]; Fig 2A). Children's speed to select a target increased over time ($\beta=$-1.24 [-1.89, -.59]; Fig 2B). Children generally produced short descriptions, and the length of descriptions tended to slightly increase over time ($\beta=$0.11 [-.03, .25]; Fig 2C).

@@

\bigskip
